### 2024-06-03 cover-letter.qmd

---
format: 
  pdf:
    include-in-header:
      text: '\pagenumbering{gobble}'

editor: visual
---

\today

Dr. Paul Montagna, Dr. Linda Deegan, Editors-in-Chief\
Estuaries and Coasts

\vfill

Dear Dr. Montagna and Dr. Deegan,

\vspace{0.5cm}

Enclosed please find the second revision of our manuscript, titled \`Hot and fresh: evidence of climate-related suboptimal water conditions for seagrass in a large Gulf coast estuary', to be considered as an original article in Estuaries and Coasts.

We have addressed the remaining comments from Dr. Wong, as noted in our attached responses. In particular, we have removed the interaction terms from the GAMs and included appropriate statistics when describing the resulting relationships. Our main conclusions remain the same. Additional minor revisions are noted in our responses.

Again, we greatly appreciate the opportunity to publish this work in Estuaries and Coasts and are confident readers will consider it a valuable contribution. Please feel free to contact us directly should additional information be needed.

\vspace{0.5cm}

Regards,

Marcus W. Beck, Tampa Bay Estuary Program, St. Petersburg, Florida, USA, [mbeck\@tbep.org](mailto:mbeck@tbep.org)

\vfill